BivData Setup

### 2020-01-28 doc/Bivalent_Analysis.Rmd
some small changes

setwd("~./MLH1repo/")
#load Clean BivData
load(file="data/CleanBivData_1.22.20.RData")

@@

In an attempt to test for and mitigate noise based on chromosome identity/physical size, I tried to isolate bivalents which are in the top quartille for SC length from their cells.

Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents. But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.

Below are examples of plots which show the SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. There are plot pairs for each category, but they are excluded for space.

Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.

@@

I'm purposing to take all the automated biv measures above the cell's 4th quartille as a data set for 'longest' bivalents which we would predict to exclude the shortest chromsomes which would have stronger chromosome size effects.

But this data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females). But it's easier to just run another set of analysis.

@@

The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures. full data set is 
`r length(Curated_BivData$Obj.ID)`. This is the breakdown of bivalent observations by category `r data.frame(table(droplevels(real.long.bivData$category)))`.

Still very unbalanced across sexes and strains ~ merge with the manual whole cell biv measures later...
Musc males are very under-represented. MSM male only has data from 1 cell.

- Try to merge this DF with the whole.cell manual measures.

@@

@@

@@

```{r show.SC.length, echo=FALSE}
#compile data
#main figures, show both mouse and bivalent level

#Curated_BivData

#Mouse_average - BivTable?

#for now remove the 
Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count <5,]

first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(30,200))
first.SC.plot
#add boxplots for mouse level means
#add second x axis, hand.foci.count

other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)

@@

#sfsL <- ggplot(lowGEN, aes(variable, prop, fill=as.factor(Group.2))) + geom_col(position = "dodge") + #scale_fill_viridis(name="Migration Rate", discrete = TRUE) + theme(axis.text.x=element_blank()) + #xlab("Frequency in the sample") + ylab("Proportion of Total Junctions") + ggtitle("C. Generation = #2,000") + ylim(0,1)

#geom_boxplot(alpha = .2, position = "dodge")+

@@

@@

@@

#sis.co.ten.hist.poin99 <- ggplot((master_curated_BivData[1:50,]), aes(x=SisCoTen, y=chromosomeLength, color= #as.factor(hand.foci.count)  ))+
#  geom_point(aes(alpha=.5)) +facet_wrap(~ mouse)
```
-->

>>>>

setwd("~./MLH1repo/")
#load Clean BivData
load(file="data/CleanBivData_1.27.20.RData")

@@

I tried to isolate bivalents which are in the top quartille for SC length from their cells.

<!-- 
Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents. But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
--> 
Below are examples of plots of SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. Most plts excluded for space.

Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data. (the extra stats are not correclty mapped)

@@

This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).

@@

The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures. The full data set is `r length(Curated_BivData$Obj.ID)`. This is the breakdown of bivalent observations by category for this long dataset are `r data.frame(table(droplevels(real.long.bivData$category)))`.

**- Try to merge this DF with the whole.cell manual measures.**

@@

new outline

1. illustrate problem(affects mostly SC length)

2. Expected impact on sex comparisons, extimated effect size of the X

3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons

4. 19 female, 19male, 20female 20 male

@@

female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]

@@

Brief lit review for SC length and MLH1 expections.

```{r show.SC.length, echo=FALSE}
#for now remove the 
#Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count #<5,]
#CHECK THAT THE ABOVE TOO MANY FOCI BIVALENTS DON"T EXSIS

#add boxplots for mouse level means
#add second x axis, hand.foci.count

#other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)

@@

Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

#nested boxplots
SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))
SC.length_Nested_boxplots

#+ +geom_jitter(aes(alpha=0.2))

@@

In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean -- are likely due to low data number.

For the 0 class chromosomes most all are around the same size of the 1CO distribution.

Add in the code for the plot under 2COs males and females

@@

Brief lit review for IFD / interference expectations  
**Petkov et al 2007,  CO interference underlies sex differences in RR**

"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb. However, the interference distance is the same in terms of bivalent length. We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes 
at the pachytene stage of meiosis."

Chrm1 genetic map,

(other human and mice refs

Tease n Hulten 2004: -- no difference between MLH1 foci in males and females

DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns

))

*Do I believe the physical/SC results?*

@@

#sis.co.ten.hist.poin99 <- ggplot((master_curated_BivData[1:50,]), aes(x=SisCoTen, y=chromosomeLength, color= #as.factor(hand.foci.count)  ))+
#  geom_point(aes(alpha=.5)) +facet_wrap(~ mouse)
```

```{r draft plots, eval=FALSE}
first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(30,200))
```

-->

### 2020-01-28 src/BivData_Setup.Rmd
some small changes

#rename foci positions
whole.cell.data_manual$Foci1 <- whole.cell.data_manual$F1
whole.cell.data_manual$Foci2 <- whole.cell.data_manual$F2
whole.cell.data_manual$Foci3 <- whole.cell.data_manual$F3

@@

>>>>

#rename cols to match BivData
whole.cell.data_manual$Foci1 <- whole.cell.data_manual$F1
whole.cell.data_manual$Foci2 <- whole.cell.data_manual$F2
whole.cell.data_manual$Foci3 <- whole.cell.data_manual$F3
whole.cell.data_manual$Foci4 <- whole.cell.data_manual$F4  #currently Foci4 is chr
whole.cell.data_manual$chromosomeLength <- whole.cell.data_manual$SC.length

whole.cell.data_manual$PER_Foci_1 <- whole.cell.data_manual$Foci1 / whole.cell.data_manual$chromosomeLength
whole.cell.data_manual$PER_Foci_2 <- whole.cell.data_manual$Foci2 / whole.cell.data_manual$chromosomeLength
whole.cell.data_manual$PER_Foci_3 <- whole.cell.data_manual$Foci3 / whole.cell.data_manual$chromosomeLength
#whole.cell.data_manual$PER_Foci_4 <- whole.cell.data_manual$Foci4 / whole.cell.data_manual$chromosomeLength

@@

#PER Foci

```{r clean.whole.cell, echo = FALSE}

#CHECKED FOR MISSING bivs -- also check for wrong directions, ect

#negative IFD
#5aug17_14jul17_LEW_f1_sp1_11_rev_blob4

#PER foci positions greater th
